---
Developing Data Products - Week 3 Assignment
Introduction
This project was created as part of the Developing Data Products course of the Coursera Data Science Specialisation.

The goal of the project is to create a web page presentation using R Markdown that features a plot created with Plotly, and to host the resulting web page on either GitHub Pages, RPubs, or NeoCities.

The interactive plot on the next slide represents the number of road accidents in Great Britain from 2005 to 2015, grouped by severity (slight, serious, or fatal).

```{r library}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

rm(list=ls())
library(plotly)
library(data.table)
library(tidyr)
library(lubridate)
library(zoo)
```
Extractdata

```{r Extractdata}
library(readr)
accidents0514 <- read_csv("~/Desktop/Coursera/Developing Data Products/Stats19_Data_2005-2014/Accidents0514.csv")

```


```{r Accidents}
accidents0514 <- accidents0514 %>%
    select(Accident_Severity, Date)

accidents15 <- read_csv("~/Desktop/Coursera/Developing Data Products/Stats19_Data_2005-2014/Accidents0514.csv")
accidents15 <- accidents15 %>%
    select(Accident_Severity, Date)
    
    
# concatenate data tables and free up environment
accidents <- rbind(accidents0514, accidents15)
rm(list = c("accidents0514", "accidents15"))
```

```{r SeverityFactor}
# convert severity to factor and add labels
accidents$Accident_Severity <- factor(accidents$Accident_Severity, levels = 1:3, labels = c("Fatal", "Serious", "Slight"))

# convert date strings to Date objects
accidents$Date <- dmy(accidents$Date)

# group data by date and severity, get count, one row per date
accident_count <- accidents %>%
    group_by(Date, Accident_Severity) %>%
    summarise(count = n()) %>%
    spread(key = Accident_Severity, value = count) %>% 
    as.data.frame()

# create a smoother for each severity to visualise general trends
loess_slight <- loess(Slight ~ as.numeric(Date), data = accident_count)
loess_serious <- loess(Serious ~ as.numeric(Date), data = accident_count)
loess_fatal <- loess(Fatal ~ as.numeric(Date), data = accident_count)
```

Road accidents in Great Britain (2005-2015)
```{r plotdata}
# plot data
plot_ly(accident_count) %>%
    add_trace(x = ~Date, y = ~Slight, type="scatter", mode = "markers", 
              name = "slight", legendgroup = "slight", 
              marker = list(color = "#52A9BD")) %>%
    add_trace(x = ~Date, y = ~Serious, type="scatter", mode = "markers",
              name = "serious", legendgroup = "serious", 
              marker = list(color = "#FFF16B")) %>%
    add_trace(x = ~Date, y = ~Fatal, type="scatter", mode = "markers",
              name = "fatal", legendgroup = "fatal", 
              marker = list(color = "#F5677D")) %>%
    add_trace(x = as.Date(loess_slight$x), y = fitted(loess_slight),
              type="scatter", mode = "lines",
              line = list(color = '#1A7A90'), 
              name = "slight Loess smoother", legendgroup = "slight", 
              hoverinfo = 'none', showlegend = FALSE) %>%
    add_trace(x = as.Date(loess_serious$x), y = fitted(loess_serious),
              type="scatter", mode = "lines",
              line = list(color = '#E9D625'),
              name = "serious Loess smoother", legendgroup = "serious",
              hoverinfo = 'none', showlegend = FALSE) %>%
    add_lines(x = as.Date(loess_fatal$x), y = fitted(loess_fatal),
              type="scatter", mode = "lines",
              line = list(color = '#DC2340'),
              name = "fatal Loess smoother", legendgroup = "fatal",
              hoverinfo = 'none', showlegend = FALSE) %>%
    layout(xaxis = list(title = "Year"),
           yaxis = list(title = "Number of Accidents")
    )
```